---
title: 'Comparing Groups: Tables and Visualizations'
author: "Bruno Flores"
date: "27 de junho de 2015"
output: html_document
---

# Simulating Consumer Segment Data

## Segment Data Definition

Separation between data definition and procedural code is a good programming 
practice:

```{r}
segVars <- c("age", "gender", "income", "kids", "ownHome", "subscribe")
segVarType <- c("norm", "binom", "norm", "pois", "binom", "binom")
segNames <- c("Suburb mix", "Urban hip", "Travelers", "Moving up")
segSize <- c(100, 50, 80, 70)
segMeans <- matrix(c(40, 0.5, 55000, 2, 0.5, 0.1,
                     24, 0.7, 21000, 1, 0.2, 0.2,
                     58, 0.5, 64000, 0, 0.7, 0.05,
                     36, 0.3, 52000, 2, 0.3, 0.2), 
                   ncol = length(segVars), byrow = TRUE)
segSDs <- matrix(c(5, NA, 12000, NA, NA, NA,
                   2, NA, 5000, NA, NA, NA,
                   8, NA, 21000, NA, NA, NA,
                   4, NA, 10000, NA, NA, NA), 
                 ncol = length(segVars), byrow = TRUE)
```

## Final Segment Data Generation

```{r}
seg.df <- NULL
set.seed(02554)
for (i in seq_along(segNames)){
        cat(i, segNames[i], "\n")
        
        this.seg <- data.frame(matrix(NA, 
                                      nrow = segSize[i], 
                                      ncol = length(segVars)))
        
        for (j in seq_along(segVars)) {
                if (segVarType[j] == "norm") {
                        this.seg[ , j] <- rnorm(segSize[i], 
                                                mean = segMeans[i, j], 
                                                sd = segSDs[i, j])
                } else if (segVarType[j] == "pois") {
                        this.seg[ , j] <- rpois(segSize[i],
                                                lambda = segMeans[i, j])
                } else if (segVarType[j] == "binom") {
                        this.seg[ , j] <- rbinom(segSize[i],
                                                 size = 1,
                                                 prob = segMeans[i, j])
                } else {
                        stop("Bad segment data type: ", segVarType[j])
                }
        }
        
        seg.df <- rbind(seg.df, this.seg)
}
```

```{r}
names(seg.df) <- segVars
seg.df$Segment <- factor(rep(segNames, times = segSize))
seg.df$ownHome <- factor(seg.df$ownHome, labels = c("ownNo", "ownYes"))
seg.df$gender <- factor(seg.df$gender, labels = c("Female", "Male"))
seg.df$subscribe <- factor(seg.df$subscribe, labels = c("subNo", "subYes"))
```

```{r}
summary(seg.df)
```

## Basic Formula Syntax

```{r}
aggregate(income ~ Segment, data = seg.df, mean)
```

A common task in marketing is cross-tabulating, separating customers into 
groups according to two (or more) factors:

```{r}
aggregate(income ~ Segment + ownHome, data = seg.df, mean)
```

## Visualization by Group: Frequencies and Proportions

```{r}
library(lattice)
histogram(~subscribe | Segment, data = seg.df)
```

The default in *histogram()* is to plot *proportions*. If we wanted the 
actual *counts*:

```{r}
histogram(~subscribe | Segment, data = seg.df, type = "count",
          layout = c(4, 1), col = c("burlywood", "darkolivegreen"))
```

Condition on more than one factor:

```{r}
histogram(~subscribe | Segment + ownHome, data = seg.df)
```

## Visualization by Group: Continuous Data

```{r}
boxplot(income ~ Segment, data = seg.df, yaxt = "n", ylab = "Income ($k)")
ax.seq <- seq(from = 0, to = 120000, by = 20000)
axis(side = 2, at = ax.seq, labels = paste(ax.seq/1000, "k", sep = ""), 
     las = 1)
```

```{r}
bwplot(Segment ~ income | ownHome, data = seg.df, horizontal = TRUE,
       xlab = "Income")
```

