---
title: "An Overview of the R Language"
author: "Bruno Flores"
date: "2 de maio de 2015"
output: html_document
---

```{r echo=FALSE}
library('ProjectTemplate')
load.project()
```

## Quick tour of R's Capabilities

Load a data ser from the book's website:
```{r}
# satData <- read.csv("http://goo.gl/UDv12g")
satData$Segment <- factor(satData$Segment)
head(satData)
```

Plot the correlation matrix (omitting the Segment column):
```{r}
# library(corrplot)
corrplot.mixed(cor(satData[,-3]))
```

Mean satisfaction for each segment:
```{r}
aggregate(iProdSAT ~ Segment, satData, mean)
```

Are the differences statistically significant:
```{r}
sat.anova <- aov(iProdSAT ~ -1 + Segment, satData)
summary(sat.anova)
```

Plot the ANOVA model to visualize confidence intervals for mean product satisfaction by segment:
```{r}
suppressMessages(suppressWarnings(library(multcomp)))
par(mar=c(4, 8, 4, 2))
plot(glht(sat.anova))
```
The X axis represents a Likert rating scale ranging 1-7 for product satisfaction.

### Structural equation model on the satisfaction data

Latent SATisfaction is observed as items iProdSAT and iSalesSAT. Latent likelihood to RECommend is observed as items iProdREC and i SalesREC. RECommendation varies with SATisfaction:
```{r}
satModel <- "SAT =~ iProdSAT + iSalesSAT
                REC =~ iProdREC + iSalesREC
                REC ~ SAT"
```

Fit the model with lavaan:
```{r}
suppressMessages(suppressWarnings(library(lavaan)))
sat.fit <- cfa(satModel, data=satData)
summary(sat.fit, fit.m=TRUE)
```

Visualize the structural model:
```{r}
# library(semPlot)
semPaths(sat.fit, what="est", residuals=FALSE, intercepts=FALSE, nCharNodes=9)
```

A structural model with path loadings for a model of product satisfaction and likelihood-to-recommend. Satisfaction has a strong relationship to likelihood-to-recommend: coefficient=0.76
